Free Probability Analysis for Reflect Surface aided Massive MIMO systems

### 2020-06-15 documentation/Paper/paper.tex
finishing correlation background and moving onto RIS optimization background

\chapter{Introduction}
\section{Background}\label{Background}
The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards, projected to reach even the Terahertz range by the 6G standard (CITE). This expansion is motivated by large bands of currently unused frequency
and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}. A primary hurdle towards efficient communication at these frequencies is the increased radio frequency (power) attenuation. For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
has been shown to exceed 100 dB. With increases in current wireless system capacity and reliability relying on the use of these high frequencies, mitigating or countering this power loss  is highly relevant.
\par
In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
and supplementing the transmission power using relays (CITE). Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission. (Would be nice to cite some actual costs here). Deploying this solution on a scale to counter
the attenuation levels relevant to new standards would pose impractical costs. One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves. In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
\par
The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
associated with a relay. The absence of an rf-chain in the RIS offers additional benefits over relays. First, reflection does not introduce the noise associated with the rf-chain. Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain). Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay. 
The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (? Does this hold if we remove the ability to choose the angle?).
For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
Example of how these surfaces are implemented (Cite).
Discuss how it has be looked at from MISO and MIMO perspective.
\par
Now discuss all of the previous work looking at the optimization of the phases. 
General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.
Need to mention anyone who has considered correlation at end in order to transition. 
\par
Early investigation leading to the great interest in MIMO systems showed the key property of capacity in an $N \times N$ MIMO system scaling linearly in $N$ (teletar). While this scaling applies to many models for the channel (cite) one key property that can impact this scaling is correlation between channels of a MIMO system.
In order to evaluate MIMO system performance in a real-world application, correlation between between the channels must be accounted for. In test data collected for (parameters) relevant to 3G and 4G wireless correlation between channels was shown to be as high as $0.5$ for a point-to-point, $4 \times 4$ system \cite{martin2000multiple} (corresponding capacity drop?). (Is this a bigger problem for higher frequencies?). Correlations cause reduction in the capacity of the system and often make the calculation of the capacity more difficult. However, with the possibility for arrays with (x) co-located antenna and RIS with (y) individually controlled reflective elements, the impact of channel correlation on system performance must be well understood before deployment. 
\par
Detail all of the work looking at correlated MIMO. 
1. Looking at correlation for finite size systems.
Explain why calculating capacity (even without correlation) for finite case is difficult/cumbersome. Describe early results for MIMO.
2. Extending to asymptotic system sizes. 
One method that has been shown to reduce the complexity of this calculation is the assumption that $N \rightarrow \infty$. This method has been used 
extensively since. 
Now list of the different types of channels/methods that have been used here. 
\cite{loyka2001channel}
\par
From the above, it should be clear that investigating the RIS  system for 
asymptotically large MIMO systems with correlation, an topic not yet explicitly investigated to the author's knowledge, is relevant to the design of upcoming wireless systems.

@@

If $\mHt$ is not known at the transmitter, the transmit power can be allocated
equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
\begin{equation}\label{no_csi_capacity}
\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
\end{equation}
where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
Expanding the expectation gives the expression
\begin{equation}\label{no_csi_capacity_aed}
\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
\end{equation}
in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.

>>>>

\chapter{Introduction}
\section{Background}\label{Background}
The spectrum of electromagnetic frequencies used by wireless communication has expanded to progressively higher frequencies with new standards, projected to reach even the Terahertz range by the 6G standard (CITE). This expansion is motivated by large bands of currently unused frequency
and the possibility of decreased antenna size (allowing for increased antenna density) \cite{akyildiz2018combating}. A primary hurdle towards efficient communication at these frequencies is the increased radio frequency (power) attenuation. For frequency bands relevant for upcoming wireless standards , freq power attenuation due to path loss  molecular absorption loss
has been shown to exceed 100 dB. With increases in current wireless system capacity and reliability relying on the use of these high frequencies, mitigating or countering this power loss  is highly relevant.
\par
In past generations of wireless communication, this problem of (power) attenuation has been alleviated by reducing the effective transmission distance 
and supplementing the transmission power using relays (CITE). Two key costs to deploying relays are the radio frequency chains used for receiving and transmitting, and the power used for transmission. (Would be nice to cite some actual costs here). Deploying this solution on a scale to counter
the attenuation levels relevant to new standards would pose impractical costs. One new solution that has attracted significant interest (check if deployed yet) are surfaces made of many elements with favorable reflective properties for high frequency electromagnetic waves (detail) and whose reflective coefficients are controlled to enable active selection of the phase of reflected waves. In this work, such a surface will generally be referred to as a reconfigurable intelligent surface (RIS).
\par
The motivating use for the RIS is to alleviate the power attenuation problem without incurring the costs of an rf-chain and transmission power 
associated with a relay. The absence of an rf-chain in the RIS offers additional benefits over relays. First, reflection does not introduce the noise associated with the rf-chain. Second, without the band-pass filters of an rf-chain, they have a frequency response over the entire spectrum (explain). Third, they are inherently full-duplex, a feature that is difficult to implement for co-located rf-chains on a relay. 
The thorough review in \cite{basar2019wireless} provides rigorous motivation of the use of RISs and a few notable points are included here.
For a single antenna, single receiver channel using an RIS with $N$ reflective elements, the power at the receiver is can be proportional to $N^2$ (? Does this hold if we remove the ability to choose the angle?).
For frequencies used in the current 5G standard, 30 GHz, an RIS with $100$ elements would have dimensions $1$ $\text{m}^2$.
(Why can we just ignore the fact that we might be able to change the reflection angle? In \cite{basar2019wireless} they introduce the idea with the ability to optimize the angle by then ignore it?).
Example of how these surfaces are implemented (Cite).
Discuss how it has be looked at from MISO and MIMO perspective.
\par
Now discuss all of the previous work looking at the optimization of the phases. 
General difficultly of the problem due to the non-convexity of the problem. Look at some different ways it's been applied.

\par
\input{correlation_review}
\par
Motivated by the potential benefit of the RIS, we now investigate for the RIS system with correlation

@@

If $\mHt$ is not known at the transmitter, the transmit power can be allocated
equally such that $\mQ_x = \mathbf{I}$. This allows for simplifying equation \eqref{capacity} to 
\begin{equation}\label{no_csi_capacity}
\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R \Expect[\Log(1 + \frac{1}{\sigma_n}\lambda_i)]
\end{equation}
where $\lambda_i$ are the eigenvalues of $\mHt \mHt^H$.
Notice that in this form, the linear scaling of capacity with N is clear.
Expanding the expectation gives the expression
\begin{equation}\label{no_csi_capacity_aed}
\mathbb{C} = \underset{\phase}{\mathop{max}} = N_R  \int_{x=0}^{\infty}\Log(1 + \frac{1}{\sigma_n}\lambda_i)p_{\lambda\lambda^H}(x) dx
\end{equation}
in which the bounds of the integral represent the range for the eigenvalues of a positive semidefinite matrix.
